Adaptive group-regularized logistic elastic net regression

### 2019-03-22 docs/manuscript_biostatistics.Rnw

\documentclass[a4paper,hidelinks]{article}
\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable,bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}

@@

@@

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands\\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge, United Kingdom \\

	\begin{abstract}
		{In high-dimensional data settings, additional information on the features is often available. Examples of such external information in omics research are: (a) $p$-values from a previous study, (b) a summary of prior information, and (c) omics annotation. The inclusion of this information in the analysis may enhance classification performance and feature selection, but is not straightforward in the standard regression setting. As a solution to this problem, we propose a group-regularized (logistic) elastic net regression method, where each penalty parameter corresponds to a group of features based on the external information. The method, termed \texttt{gren}, makes use of the Bayesian formulation of logistic elastic net regression to estimate both the model and penalty parameters in an approximate empirical-variational Bayes framework. Simulations and an application to a colon cancer microRNA study show that, if the partitioning of the features is informative, classification performance and feature selection are indeed enhanced.}
	\end{abstract}

	\noindent\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes

	\noindent\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}

	\section{Introduction}
	Prediction from high-dimensional data is a challenge common to many fields of research. Examples of such high-dimensional prediction problems arise in computer vision, stock market prediction and disease diagnosis from omics data. In this paper, we specifically focus on the latter setting. High dimensionality of data introduces several issues in the estimation of prediction models, especially: (a) unidentifiable models, (b) highly correlated predictor variables and (c) non-trivial variable selection.

@@

	In omics research, sources of such prior knowledge on the variables are often available. The information can, for example, come in the form of (a) results on the same molecular features obtained in a previous study (e.g. $p$-values), (b) information from a publicly available database that summarizes the prior information on the molecular features involved (e.g., the Cancer Gene Census \cite[]{futreal_census_2004}), and (c) omics annotation (e.g. the location of a gene on the chromosome). Although this information can rarely be directly included in the statistical analysis, it may still be useful and informative for the study at hand.

	An example of such a study with extra information available is investigated in Section \ref{sec:app}. INTRODUCE DATA. This problem is not straightforward to solve with the standard methods available in literature. At first sight, such an \textit{a priori} grouping of the features suggests the group lasso \cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which penalises and selects features at the group level. There are two limitations to the group lasso: (a) it selects groups as a whole instead of single features and (b) it is not adaptive: the amount of penalization a group receives depends on a `global' penalty parameter common to all groups. Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).

	A natural way to deal with (b) is through differential penalization: that is, each group of variables receives its own penalty parameter. An apparent issue with this differential penalization is the estimation of the penalty parameters. Naive estimation may be done by cross-validation. However, cross-validation requires re-estimation of the model over a grid of penalty parameters. The size of this grid increases exponentially with the number of penalty parameters. Consequently, it quickly becomes computationally infeasible. We therefore propose an efficient alternative: empirical Bayes estimation of the penalty parameters, which corresponds to hyperparameter estimation in the Bayesian prior framework. Because of the ubiquity of binary outcome data in omics research, we focus on the logistic elastic net. Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in spirit. Nevertheless, one clear advantage of our method over VBVS is the possibility of adaptive inclusion of external information on the features to aid in prediction and feature selection.

	In line with the above, our proposed method is (a) subjective: the few remaining tuning parameters may be cross-validated, (b) adaptive: prior information is automatically and subjectively weighed with respect to its informativeness, (c) fast compared to full Bayesian analysis or cross-validation, and (d) requires minimal input by the user: only the data and grouping of the features is required.

	We introduce the Bayesian logistic (generalised) elastic net model in Section \ref{sec:model}. In Section \ref{seq:estimation} we derive a variational approximation to this model and use this novel approximation in the empirical Bayes estimation of multiple, group-specific penalty parameters. In Sections \ref{sec:simulations}, and \ref{sec:app}, we demonstrate the approach in a simulation study and a real data example, respectively. In Section \ref{sec:discussion} we conclude by both discussing some differences and parallels between the proposed method, termed \texttt{gren}, and other methods in the literature.

	\section{Model}\label{sec:model}
	\subsection{Logistic elastic net regression}\label{sec:logitmodel}
	In logistic regression the outcome variables are assumed to be binary or sums of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l \in \{0,1\}$ for $i=1, \dots, n$). The binomial logistic model relates the responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix} x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} \( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is the binomial distribution with number of trials $m$ and probability $\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + \exp(\x\tr_i \bbeta)]$. Note that if $m_i=1$ for $i=1, \dots, n$, the model reduces to a binary logistic regression model. Throughout the rest of the paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots & \x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.

	In elastic net regression, the penalised likelihood is maximised to yield parameter estimates:
	$$
	\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \norm{\bbeta}_1 - \frac{\lambda_2}{2} \norm{\bbeta}_2^2,
	$$
	with $\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ and $\y = \begin{bmatrix} y_1 & \cdots & y_n \end{bmatrix} \tr$. The likelihood term is thus complemented with a penalty term consisting of the $L_1$ and $L_2$-norm of the parameters, multiplied by respective penalty parameters $\lambda_1$ and $\lambda_2$. We have scaled the penalty parameters by $\frac{1}{2}$, to make them correspond with the Bayesian version in the following. The maximiser of the penalised likelihood corresponds to the posterior mode of a Bayesian elastic model with prior \cite[]{li_bayesian_2010,zou_regularization_2005}:
	\begin{equation}\label{eq:elasticnetprior}
	\beta_j \sim g(\lambda_1,\lambda_2) \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta_j| + \lambda_2 \beta_j^2)\].
	\end{equation}
	Here, $g(\lambda_1,\lambda_2)$ denotes a normalizing constant that is dependent on $\lambda_1$ and $\lambda_2$ (given in Section 2 of the Supplementary Material (SM), along with further details on the elastic net prior). The elastic net combines $L_1$ and $L_2$-norm penalisation such that the model parameters are shrunken towards zero. The $L_1$-norm may set some of the estimates exactly to zero, thus automatically selecting features. The $L_2$-norm ensures that correlated features behave similarly, as one would generally require. We extend the elastic net model to the generalised elastic net to allow for different penalty parameters per group of features.

	\subsection{The generalised elastic net}\label{sec:genen}
	Assume we have a partitioning of the features in $G$ groups, such that each feature belongs to one group. Let $\G(g)$ be the feature index set of group $g$ for $g=1, \dots, G$. To apply differential penalisation we include a group-specific weight $w_g \in \mathbb{R}_{>0}$. The model parameters are now estimated by:
	\begin{align}
	\hat{\bbeta} & := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 \label{eq:weightedlik}\\ 
	& = \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, \nonumber
	\end{align}
	where we wrote $w^2_g = \lambda'_g$ to emphasise that these group-specific weights may be interpreted as penalty multipliers. Throughout the following we assume that the geometric mean of the multipliers, weighted by their respective group sizes, is one, such that the average shrinkage of the model parameters is determined by the `global' $\lambda_1$ and $\lambda_2$. That is, we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G (\lambda'_g)^{|\G(g)|}=1$. The multiplier appears in square root form in the $L_1$-norm term to ensure that penalisation on the parameter level is the same for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in (\ref{eq:weightedlik}).

	\cite{li_bayesian_2010} show that (\ref{eq:elasticnetprior}) may be written as a computationally more convenient scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$. Using this result, we write the generalised elastic net model in its Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i \bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this Bayesian formulation the penalty parameters in $\blambda = \begin{bmatrix} \lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ play the role of the hyperparameters in a Bayesian hierarchical model.

	\section{Estimation}\label{seq:estimation}
	\subsection{Empirical Bayes}\label{seq:empiricalbayes}
	If the penalty parameters are known, estimation of the frequentist elastic net model parameters or finding the posterior of the generalized elastic net model is feasible with small adjustments of the available algorithms \cite[]{friedman_regularization_2010,zou_regularization_2005} or MCMC samplers \cite[]{li_bayesian_2010}. Determining these penalty parameters, however, is not straightforward.

@@

	\subsection{Variational Bayes}\label{sec:variationalbayes}
	Variational Bayes is a widely used method to approximate Bayesian posteriors. It has successfully been applied in a wide range of applications, including genetic association studies \cite[]{carbonetto_scalable_2012} and gene network reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior is approximated by a tractable form and estimated by optimizing a lower bound on the marginal likelihood of this model (see Section 3 of the SM for the lower bound of the proposed model). For an extensive introduction and concise review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, respectively.

	To simplify the computations of our variational approximation, we follow \cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for $i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of the $y_i$ and P\'{o}lya-Gamma distributed (see Section 4 of the SM for more details). We augment Model (\ref{eq:enmodel2}) with:
	\begin{equation}\label{eq:logmodel}
	\bomega | \bbeta \sim \prod_{i=1}^n \mathcal{PG}\(m_i, |\x_i \tr \bbeta| \).
	\end{equation}
	Our variational Bayes approximation to the posterior distribution of (\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of parameters. We choose the blocks such that:
	\begin{equation}\label{eq:varbayesapproximation}
	p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
	\end{equation}
	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and $\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, calculus of variations gives the optimal distributions $q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} [\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the Kullback-Leibler divergence of the posterior to the approximate distribution \cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation}) renders both the posterior parameter calculations and the expected joint log likelihood as introduced in (\ref{eq:ebEM}) tractable.

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal distributions in our variational Bayes implementation for the model parameters as:
	\begin{equation}\label{eq:varbayesdistr}
	q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)\text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\), 
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian distribution (See SM Section 5 for the derivations). The so-called variational parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we update them by:
	\begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		\bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=\diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		\bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\ 
		c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		\chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries $\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & \cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and $\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th element of a column vector, respectively. Naive calculation of the variational parameters is computationally expensive. In Section 6 of the SM we show that informed calculation of the parameters results in a significant reduction of computational complexity.

>>>>

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
@

% read figure code
<<figures, include=FALSE>>=
read_chunk("code/figures.R")
@

\documentclass[oupdraft]{bio}
\usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable}

@@

	% Title of paper
	\title{Adaptive group-regularized logistic elastic net regression}

	% List of authors, with corresponding author marked by asterisk
	\author{MAGNUS M. M\"UNCH$^{\ast 1,2}$, CAREL F.W. PEETERS$^{1}$, AAD W. VAN 
	DER VAART$^{2}$, \\* MARK A. VAN DE WIEL$^{1,3}$\\[4pt]
	% Author addresses
	\textit{1. Department of Epidemiology \& Biostatistics, Amsterdam Public 
	Health research institute, Amsterdam University medical centers, PO Box 7057,
	1007 MB Amsterdam, The Netherlands \\*
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom}
	\\[2pt]
	% E-mail address for correspondence
	{m.munch@vumc.nl}}

	% Running headers of paper:
	\markboth%
	% First field is the short list of authors
	{M. M. M\"unch and others}
	% Second field is the short title of the paper
	{Adaptive group-regularized logistic elastic net regression}

@@

	% Add a footnote for the corresponding author if one has been
	% identified in the author list
	\footnotetext{To whom correspondence should be addressed.}

	\begin{abstract}
		{In high-dimensional data settings, additional information on the features 
		is often available. Examples of such external information in omics research
		are: (a) $p$-values from a previous study, (b) a summary of prior 
		information, and (c) omics annotation. The inclusion of this information in
		the analysis may enhance classification performance and feature selection,
		but is not straightforward in the standard regression setting. As a solution
		to this problem, we propose a group-regularized (logistic) elastic net
		regression method, where each penalty parameter corresponds to a group of
		features based on the external information. The method, termed 
		\texttt{gren}, makes use of the Bayesian formulation of logistic elastic net
		regression to estimate both the model and penalty parameters in an 
		approximate empirical-variational Bayes framework. Simulations and
		applications in two different molecular marker and cancer type studies show
		that, if the partitioning of the features is informative, classification 
		performance and feature
		selection are indeed enhanced.}
		{Empirical Bayes; High-dimensional data; Prediction; Variational Bayes}
	\end{abstract}

	\section{Introduction}\label{sec:introduction}
	Prediction from high-dimensional data is a challenge common to many fields of
	research. Examples of such high-dimensional prediction problems arise in
	computer vision, stock market prediction and disease diagnosis from omics 
	data. In this paper, we specifically focus on the latter setting. High
	dimensionality of data introduces several issues in the estimation of 
	prediction models, especially: (a) unidentifiable models, (b) highly 
	correlated predictor variables and (c) non-trivial variable selection.

@@

	In omics research, sources of such prior knowledge on the variables are often
	available. The information can, for example, come in the form of (a) results 
	on the same molecular features obtained in a previous study (e.g. $p$-values),
	(b) information from a publicly available database that summarizes the prior
	information on the molecular features involved (e.g., the Cancer Gene Census
	\cite[]{futreal_census_2004}), and (c) omics annotation (e.g. a gene's 
	chromosome location). Although this information can rarely be directly
	included in the statistical analysis, it may still be useful and informative
	for the study at hand.

	Two such studies are introduced in Sections \ref{sec:colorectal} and
	\ref{sec:lymphnode}. The first study is concerned with treatment 
	response prediction for colorectal cancer patients, based on sequenced
	microRNAs. In the second study we attempt to predict lymph node metastases,
	using RNAseq data. In both settings additional information on the 
	features is available in the form of a partitioning. Furthermore, in both
	examples data is sparse, so any extra included information may help to boost
	predictive performance. Inclusion of such feature partitions in 
	the analysis is not straightforward with standard methods.

	At first sight, such an \textit{a priori} grouping of the features suggests 
	the group lasso
	\cite[]{meier_group_2008, yuan_model_2006} or one of its extensions, because
	the group lasso penalises and selects features at the group level. 
	There are two limitations to this approach: the group lasso (a) selects groups
	as a whole instead of single 
	features and (b) does not penalise adaptively: all groups are penalised 
	equally. Extensions such as the sparse group lasso 
	\cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).

	A natural way to deal with (b) is through differential penalization: that is,
	each group of variables receives its own penalty parameter. An apparent issue
	with this differential penalization is the estimation of the penalty 
	parameters. Naive estimation may be done by cross-validation. However, 
	cross-validation requires re-estimation of the model over a grid of penalty
	parameters. The size of this grid increases exponentially with the number of
	penalty parameters. Consequently, it quickly becomes computationally 
	infeasible. We therefore propose an efficient alternative: empirical Bayes
	estimation of the penalty parameters, which corresponds to hyperparameter
	estimation in the Bayesian prior framework. Because of the ubiquity of binary
	outcome data in omics research, we focus on the logistic elastic net.

	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
	spirit. Nevertheless, one clear advantage of our method over VBVS is the
	possibility of adaptive inclusion of external information on the features to
	aid in prediction and feature selection. Bayesian versions of support vector 
	machines have been used in classification problems as well 
	\cite[]{chakraborty_bayesian_2011}, but these methods also lack 
	adaptive inclusion of external information on the features.

	In line with the above, our proposed method is (a) subjective: the 
	hyper-parameters are estimated from the data and not chosen on objective 
	grounds, (b) adaptive: prior information is automatically and subjectively 
	weighed with respect to its informativeness, (c) fast compared to full 
	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
	user: only the data and grouping of the features is required. We emphasize that 
	our method is different from recent developments in grouped penalized 
	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
	features, the observations are partitioned.

	The rest of the paper is structured as follows. We introduce the Bayesian 
	logistic generalised elastic net model in Section
	\ref{sec:model}. In Section \ref{sec:estimation} we derive a variational
	approximation to this model and use this novel approximation in the empirical
	Bayes estimation of multiple, group-specific penalty parameters. In Section
	\ref{sec:simulations} we compare the method in a simulation study. In Sections
	\ref{sec:colorectal} and \ref{sec:lymphnode}, we demonstrate the approach in 
	two data sets. In Section \ref{sec:discussion} 
	we conclude by both discussing some differences and
	parallels between the proposed method, termed \texttt{gren}, and other methods
	in the literature.

	\section{Model}\label{sec:model}
	In logistic regression the outcome variables are assumed to be binary or sums
	of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l 
	\in \{0,1\}$ for $i=1, \dots, n$). The binomial logistic model relates the
	responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix}
	x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} 
	\( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is 
	the binomial distribution with number of trials $m$ and probability 
	$\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + 
	\exp(\x\tr_i \bbeta)]$. Note that if $m_i=1$ for $i=1, \dots, n$, the model
	reduces to a binary logistic regression model. Throughout the rest of the 
	paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots &
	\x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n
	x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.

	Assume we have a partitioning of the features in $G$ groups, such that each
	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
	$g$ for $g=1, \dots, G$ and let $w_g \in \mathbb{R}_{>0}$ denote a 
	group-specific weight. In a generalised elastic net regression, the penalised 
	likelihood is maximised to yield parameter estimates:
	\begin{align}
	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) - 
	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 
	  \label{eq:weightedlik}\\
	  & = \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  \frac{\lambda_1}{2} 
	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
	  \nonumber
	\end{align}
	where $\L(\y ; \bbeta)$ denotes the likelihood function of $\bbeta$ given the 
	observed data $\y = \begin{bmatrix} 
	y_1 & \cdots & y_n \end{bmatrix} \tr$, and 
	$\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ are the `global' penalty 
	parameters. We wrote $w^2_g = \lambda'_g$ to emphasise that these 
	group-specific weights may be interpreted as penalty multipliers. Note that
	the regular elastic net is recovered by setting $\forall g: \lambda'_g=1$.

  Throughout the following we assume that the geometric mean of the multipliers, 
	weighted by their 
	respective group sizes, is one, such that the average shrinkage of the model
	parameters is determined by $\lambda_1$ and $\lambda_2$. That is,
	we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G 
	(\lambda'_g)^{|\G(g)|}=1$. The multiplier appears in square root form in the
	$L_1$-norm term to ensure that penalisation on the parameter level is the same
	for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in
	(\ref{eq:weightedlik}). The elastic net combines $L_1$ and $L_2$-norm 
	penalisation such that the model parameters are shrunken towards zero. 
	The $L_1$-norm may set some of the estimates exactly to zero, thus 
	automatically selecting features. The $L_2$-norm ensures that correlated 
	features behave similarly, as one would generally require.

	The maximiser of the penalised likelihood in the elastic net, corresponds 
	to the posterior mode of a 
	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
	\cite{li_bayesian_2010} show that the elastic net prior 
	(see Supplementary Material (SM) Section \ref{sm-sec:enetprior} for 
	details on this prior) may be written as
	a computationally more convenient scale mixture of normals, with mixing 
	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
	\tr$. Using this result, we write the generalised elastic net model in its 
	Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
		\bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
		}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
	play the role of the hyperparameters in a Bayesian hierarchical model.

	\section{Estimation}\label{sec:estimation}
	\subsection{Empirical Bayes}\label{sec:empiricalbayes}
	If the penalty parameters are known, estimation of the frequentist elastic net
	model parameters or finding the posterior of the generalized elastic net model
	is feasible with small adjustments of the available algorithms 
	\cite[]{friedman_regularization_2010,zou_regularization_2005} or MCMC samplers
	\cite[]{li_bayesian_2010}. Determining these penalty parameters, however, is 
	not straightforward.

@@

	\subsection{Variational Bayes}\label{sec:variationalbayes}
	Variational Bayes is a widely used method to approximate Bayesian posteriors. 
	It has successfully been applied in a wide range of applications, including 
	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
	reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior 
	is approximated by a tractable form and estimated by optimizing a lower bound 
	on the marginal likelihood of this model (see Section 
	\ref{sm-sec:objectivefunction} of the SM for the 
	lower bound of the proposed model). For an extensive introduction and concise 
	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, 
	respectively.

	To simplify the computations of our variational approximation, we follow
	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 
	\ref{sm-sec:polyagamma} of the SM for more 
	details). We augment Model (\ref{eq:enmodel2}) with:
	\begin{equation}\label{eq:logmodel}
	  \bomega | \bbeta \sim \prod_{i=1}^n \mathcal{PG}\(m_i, |\x_i \tr \bbeta| \).
	\end{equation}
	Our variational Bayes approximation to the posterior distribution of 
	(\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of 
	parameters. We choose the blocks such that:
	\begin{equation}\label{eq:varbayesapproximation}
	  p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) 
	  = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
	\end{equation}
	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and 
	$\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, 
	calculus of variations gives the optimal distributions 
	$q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} 
	[\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the 
	Kullback-Leibler divergence of the posterior to the approximate distribution 
	\cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation})
	renders both the posterior parameter calculations and the expected joint log 
	likelihood as introduced in (\ref{eq:ebEM}) tractable.

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
	distributions in our variational Bayes implementation for the model parameters
	as:
	\begin{equation}\label{eq:varbayesdistr}
	  q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } 
	  q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)
	  \text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} 
	  \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
	distribution (See SM Section \ref{sm-sec:variationaldistributions} 
	for the derivations). The so-called variational
	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
	update them by:
	\begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
	element of a column vector, respectively. Naive calculation of the variational
	parameters is computationally expensive. In Section 
	\ref{sm-sec:computationalcomplexity} of the SM we show that 
	informed calculation of the parameters results in a significant reduction of 
	computational complexity.
